Category-based and location-based volitional covert attention affect memory at different timescales

### 2023-04-10 paper/main.tex
typo fix in abstract

Our ongoing subjective experiences, and our memories of those experiences, are
shaped by our prior experiences, goals, and situational understanding. These
factors influence how we allocate our attentional resources over different
aspects of our ongoing experiences. These attentional shifts may happen overtly
(e.g., when we change where we are looking) or covertly (e.g., without any
external physical manifestation). Additionally, we may attend to what is
happening at a specific spatial location (e.g., because we think something
important is happening there) or we may attend to particular features
irrespective of their locations (e.g., when we search for a friend's face in a
crowd versus a desired item in a grocery store). We ran a covert attention
experiment with two conditions that asked participants to either sustain or
vary the focus of the categories and locations they were attending. Later, the
participants performed a recognition memory task for attended, unattended, and
novel stimuli. Participants in both conditions recognized attended stimuli more
readily than unattended or novel stimuli. Participants in the ``sustained
attention'' condition also exhibited a recognition advantage for
attended-category images from the unattended location, whereas participants in
the ``variable attention'' condition exhibited a recognizing advantage for
attended-location images from the unattended category. Our findings suggest
that covert attention enhances memory encoding of attended stimuli, and that
different aspects of partial attention affect memory encoding over different
timescales.

>>>>

Our ongoing subjective experiences, and our memories of those experiences, are
shaped by our prior experiences, goals, and situational understanding. These
factors influence how we allocate our attentional resources over different
aspects of our ongoing experiences. These attentional shifts may happen overtly
(e.g., when we change where we are looking) or covertly (e.g., without any
external physical manifestation). Additionally, we may attend to what is
happening at a specific spatial location (e.g., because we think something
important is happening there) or we may attend to particular features
irrespective of their locations (e.g., when we search for a friend's face in a
crowd versus a desired item in a grocery store). We ran a covert attention
experiment with two conditions that asked participants to either sustain or
vary the focus of the categories and locations they were attending. Later, the
participants performed a recognition memory task for attended, unattended, and
novel stimuli. Participants in both conditions recognized attended stimuli more
readily than unattended or novel stimuli. Participants in the ``sustained
attention'' condition also exhibited a recognition advantage for
attended-category images from the unattended location, whereas participants in
the ``variable attention'' condition exhibited a recognition advantage for
attended-location images from the unattended category. Our findings suggest
that covert attention enhances memory encoding of attended stimuli, and that
different aspects of partial attention affect memory encoding over different
timescales.
